NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations

### 2022-12-21 MANUSCRIPT/manuscript.tex
Major update to the manuscript with new simulation data, plots, and text. Mainly by Matti Javanainen.

%% ****** Start of file aiptemplate.tex ****** %
%%
%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
%%   Version 4.1 of 9 October 2009.
%%
%
% This is a template for producing documents for use with 
% the REVTEX 4.1 document class and the AIP substyles.
% 
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.

%\documentclass[aip,graphicx]{revtex4-1}
%\documentclass[aip,reprint]{revtex4-1}

%\usepackage{graphicx}

%\draft % marks overfull lines with a black rule on the right
%\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}

\documentclass[aps,prl,superscriptaddress,twocolumn]{revtex4}

%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}
\usepackage{xr}
\externaldocument{manuscriptSUPPL}
\begin{document}

% Use the \preprint command to place your local institutional report number 
% on the title page in preprint mode.
% Multiple \preprint commands are allowed.
%\preprint{}

\title{NMRlipids III: Lipid-cholesterol interactions in atomistic resolution molecular dynamics simulations} %Title of paper

% repeat the \author .. \affiliation  etc. as needed
% \email, \thanks, \homepage, \altaffiliation all apply to the current author.
% Explanatory text should go in the []'s, 
% actual e-mail address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro for the type of information

% \affiliation command applies to all authors since the last \affiliation command. 
% The \affiliation command should follow the other information.

\author{Fernando Favela-Rosales}
\affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}

\author{Peter Heftberger}
%\email[]{samuli.ollila@aalto.fi}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
\affiliation{University of Graz}

\author{Matti Javanainen}
\affiliation{Department of Physics, Tampere University of Technology, Tampere, Finland}
\affiliation{University of Helsinki}

\author{Jesper J. Madsen}
\affiliation{Department of Global Health, College of Public Health}
\affiliation{University of South Florida}

\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}

\author{Markus Miettinen}
\affiliation{MPI}

\author{O. H. Samuli Ollila}
\email[]{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}

\author{Georg Pabst}
\affiliation{University of Graz}

\author{Thomas Piggot}
\affiliation{University of Southampton}

% Collaboration name, if desired (requires use of superscriptaddress option in \documentclass). 
% \noaffiliation is required (may also be used with the \author command).
%\collaboration{}
%\noaffiliation

@@

\begin{abstract}
% insert abstract here
The quantitative quality of lipid-cholesterol interactions in atomistic resolution models will be determined against 
NMR and scattering data.
\end{abstract}

%\pacs{}% insert suggested PACS numbers in braces on next line

\maketitle %\maketitle must follow title, authors, abstract and \pacs

% Body of paper goes here. Use proper sectioning commands. 
% References should be done using the \cite, \ref, and \label commands

%\label{}
\section{Introduction}

\todo{We should emphasize the validation aspect of intermolecular interactions in binary lipid mixtures in the introduction.}

Details of intermolecular interactions determine the phase
behaviour and details lateral organization of lipid and bilayers
containing cholesterol \cite{ipsen87}.
Formation of lateral heterogeneities \cite{kinnunen91}, lipid rafts \cite{simons97}
and superlattices \cite{somerharju09} in cellular membranes have been
suggested to be driven by interactions between cholesterol
and lipids. While detailed experimental information of these interactions
is relatively sparse, atomistic resolution molecular dynamics simulations
have been widely applied to give detailed explanations of lipid bilayer
lateral organization and lipid-cholesterol interactions \cite{rog14,sodt14,??}.
However, the simulations must reproduce the measured experimental details, such as NMR
order parameters and scattering form factors, to be useful in intepretation
of molecular details in lipid bilayer mixtures.

Simulations qualitatively reproduce the cholesterol condensation effect,
but quantitative comparison to robust experimental data has not been
typically done. This is partly due to the lack of availability of
systematic experimental data sets for the comparison and partly
due to the lack of protocols for validating intermolecular interactions
in lipid bilayer simulations against experiments.
Scattering experiments are more difficult to interpret for
mixed lipid bilayers than for single component bilayers \cite{pan12,Heftberger15,Marquardt15,??},
thus area per molecule values, one of the main quantity used to
compare simulations to experiments, have not been
available for lipid cholesterol mixtures. Systematic experimental
data set for lipid C-H bond order parameters with different cholesterol
concentrations in POPC bilayer has been published only relatively recently \cite{ferreira13}.

In this work we present the experimental scattering form factor
data for POPC-cholesterol mixtures by systematically increasing the
cholesterol concentration, scanning the same concentrations that were previously studied with ssNMR~\cite{ferreira13}.
Our goal is to show that the combination
of systematically measured C--H bond order parameter and scattering form
factor data can be used to validate the quality of lipid--cholesterol
intermolecular interactions in MD simulations. MD simulations can be also
potentially used to give structural interpretation for form factors
measured from mixed systems, which is a major challenge for current methods \cite{pan12,Heftberger15,Marquardt15,??}. 
The approach should be also applicable for mixed lipid bilayers with
other than lipid--cholesterol mixtures.

%Importance of phospholipid cholesterol interactions in cellular membranes
%have been reckognized for long time \cite{??}. Various ideas about 
%lateral heterogeneities driven interactions between cholesterol, phosphatidylcholine
%other other membrane lipids has been proposed \cite{??}. 
%Especially coexistence of liquid-ordered and liquid-disordered phases
%in PC-cholesterol mixtures has been suggested to be important in
%biological lipid membranes \cite{??}.

%Phase diagram of PC lipid cholesterol systems, including lo-ld coexistence
%can be explained by simple models where cholesterol is assumed to 
%induce order in neighboring lipids \cite{??}. This is important
%achievement in order to understand biomembrane organization. 
%However, also more detailed interactions driving interactions 
%between cholesterol and other membrane lipid molecules are
%often relevant. Several attempts to understand these interactions
%have been made by using classical molecular dynamics simulations \cite{??}.
%These simulations give atomistic resolution information, which not
%directly accessible with any other techinique. However, 
%the simulations often suffer from inaccuracies in force field
%descriptions. Acyl chain region is usually well described in 
%molecular dynamics simulation models, which is not the case for 
%glycerol backbone and headgroup \cite{??}. 
%The same applies to the systems with cholesterol. Practically all simulation
%models qualitatively reproduce the condesing due to cholesterol, observed
%as increase in acyl chain order parameters and decrease of area per molecule.
%However, the some models significantly overestimate the changes in headgroup
%due to the addition of cholesterol \cite{??}.

%On the other hand, detailed analysis of lateral organization of 
%lipid-cholesterol mixtures reveal significant differences between
%different models. For example, significant angular organization
%of lipids around cholesterol was observed in simulations with Berger
%Höjtje models, while CHARMM36 simulations did not show such behaviour.

Order parameters for C-H bond vectors in lipid bilayer systems, measured
with $^{13}$C or $^{2}$H NMR techniques, give indirect information about the structural
sampling of indidual molecules \cite{ollila16}. The observed increase of
acyl chain order parameters with added cholesterol in simulations and
experiments can be explained by increased trans conformations in acyl chains \cite{ferreira13,??},
which is suggested to play critical role in the phase behaviour of PC-lipid--cholesterol
mixtures \cite{ipsen87}. Consequently, the correct cholesterol ordering effect is
expected to be a necessary condition for a model used to understand lipid--cholesterol
phase behaviour.

@@

\subsection{X-ray scattering experiments}

SAXS data on POPC multilamellar vesicles (MLVs) at various cholesterol concentrations
has been measured. Data have been obtained at the EMBL BioSAXS beamline (Hamburg) using 20 keV photons, T = 27$^\circ$C. Data were analyzed in terms of the SDP-GAP model described in Heftberger et al., J. Appl. Cryst. 2013 and Heftberger et al. Biophys. J. 2015. Data from MLVs are a convolute of structure factor (the crystalline lattice) and form factor. By fitting the scattered intensity data we obtain both contributions. Here we posted only form factors (ASCII format). For information on the quality of the fit we also give plots of the fitted intensity data. The electron density profile has been modelled in terms of the SDP model (see papers by Kucerka and coworkers), that is volume distribution functions are modelled by individual Gaussians or error functions. Cholesterol is also accounted for by two Gaussians. This model has been proposed by Jianjun Pan (USF, Tampa, FL), but is to the best of our knowledge not published (see also PhD Thesis by Peter Heftberger). Additional figures show the volume distribution functions and the resulting electron density profiles.

@@

\subsection{Molecular dynamics simulations}

Molecular dynamics simulation data was collected using the open collaboration method \cite{botan15}.
Simulated systems are listed in Table \ref{systemsCHOL} and the full simulation
details are given in references or in supplementary material.

>>>>

\documentclass[aps,prl,superscriptaddress]{revtex4-2}

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}
\usepackage{booktabs}
\usepackage{xr}

\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}

\externaldocument{MANUSCRIPT/manuscriptSUPPL}

\begin{document}

\title{NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations}

\author{Fernando Favela-Rosales}
\affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
%
\author{Peter Heftberger}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
%
\author{Matti Javanainen}
\email{matti.javanainen@helsinki.fi}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}
%
\author{Jesper J. Madsen}
\affiliation{Department of Global Health, College of Public Health}
\affiliation{University of South Florida}
%
\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
%
\author{Markus Miettinen}
\affiliation{MPI}
%
\author{O. H. Samuli Ollila}
\email[]{samuli.ollila@helsinki.fi}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}
%
\author{Georg Pabst}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
\affiliation{BioTechMed-Graz, Graz 8010, Austria}
%
\author{Thomas Piggot}
\affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
\affiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}

@@

\begin{abstract}
Cholesterol is a central building block of biomembranes, where it induces orientational order, slows down diffusion, and renders membranes generally stiffer. Molecular dynamics simulations have played a crucial role in resolving these effects and the underlying molecular picture, yet it has recently become evident that different simulation models provide quantitatively different behavior. As the field progresses towards simulations of more complex membranes, the fundamental interactions such as those between cholesterol and phospholipids---the key building blocks of biomembranes---such be modelled accurately.
%
Here, we perform a systematic comparison of the commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol. Comparison to deuterium order parameters and lateral diffusion coefficients from NMR as well as electron density profiles from X-ray scattering reveals that none of the existing force field outperforms the others across the physiologically relevant cholesterol concentrations. The Amber-compatible force fields show the smallest deviations from experiment with Slipids and Lipid17 models excelling at low and high cholesterol concentrations, respectively. The development of force fields or the refinement of existing ones can benefit from the included X-ray scattering data set.
\end{abstract}

\maketitle

\section{Introduction}

Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} which are unevenly distributed in the membrane plane and across the leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}. A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations between $\sim$10~mol-\% (endoplasmic reticulum) and up to $\sim$50~mol-\% (plasma membrane) \cite{van2008membrane}. CHOL has the unique ability to order neighbouring lipids and thus induce the liquid ordered (L\textsubscript{o}) phase in model membranes \cite{mouritsen2004s,ipsen87,kinnunen91,rog2009ordering}. In the cellular setting, the interaction between phospholipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}. This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation} \textit{via} the hydrophobic mismatch mechanism \cite{killian1998hydrophobic}, in addition to the direct modulation of protein function by bound CHOL molecules \cite{gimpl2016interaction}.

While detailed experimental information of the interactions among phospholipids and CHOL is relatively sparse, atomistic resolution molecular dynamics (MD) simulations have been widely applied to provide a detailed view of the lateral organization of lipid bilayers as well as lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed}. However, the trustworthiness of any prediction from an MD simulation relies on the quality of the used model. Fortunately, atomistic \emph{force fields} have seen rapid development recently \cite{leonard2019developing}. The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the model by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig15b}, respectively. Moreover, OPLS3e \cite{harder2016opls3} contains parameters for lipids, but CHOL is currently not included. Thanks to the compatibility with protein force fields, the lipid models can be used in studies of membrane proteins or protein--membrane interactions in addition to sole lipid membranes. Notably, simulations using CHARMM36, Lipid17, and Slipids are now readily set up using CHARMM-GUI for multiple simulation engines \cite{lee16,lee2020charmm}. Despite this ease-of-use, the choice of a force field for an MD simulation can be challenging, as each model and even more so their combinations come with their own sets of pros and cons. Unfortunately, careful comparisons of force field quality, especially for capturing the features of lipid mixtures, are scarce in the literature. This largely results from the lack of systematic, comprehensive, and quantitative datasets from both experiments and simulations.

We have previously performed such comparisons for the structure of phosphatidylcholine (PC) lipids \cite{botan15}, for PC--ion interactions \cite{catte2016molecular}, for phosphatidylserine lipids \cite{antila2019headgroup}, and for phoshatidylethanolamine and phosphatidylglycerol lipids \cite{bacle2021inverse}. Here, we extend this work to mixtures of PC and CHOL. We compare the ability of popular force fields to reproduce effects of CHOL on the structure and dynamics of a PC lipid bilayer, and thereby set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems. Moreover, if some of the studied force fields satisfactorily reproduce experimental behavior at least in some range of CHOL concentrations, the molecular level picture provided by it can be assumed to mimic reality in this concentration regime.

Our earlier work has demonstrated that the structural ensembles of lipids from an MD simulation can be directly compared to those from solid state NMR experiments \cite{botan15,catte2016molecular,antila2019headgroup,bacle2021inverse}. This approach based on order parameters is quantitative, has small error estimates, is highly reproducible, and does not suffer from perturbation of probes used in many other techniques. Thus, we compare our simulation results with different force fields to the systematic NMR data set of PC C--H bond order parameters with varying cholesterol concentrations \cite{ferreira13}. In addition, we extracted electron density profiles across the lipid bilayer from scattering form factor data measured in this work for a range of PC--CHOL mixtures using small angle X-ray scattering. These profiles can be directly compared with the same profiles calculated from the simulation data. Whereas NMR provides data on the conformations of individual lipids, form factor captures the details of the structure of the entire lipid bilayer \cite{ollila16}. One limitation in comparing simulations of lipid mixtures with experimental data is that the various experiments are performed at different concentrations, and thus mimicking all compositions in simulations is tedious. Here, we demonstrate that data extracted from well-sampled simulations can readily be interpolated to any CHOL concentration, allowing the comparison of a single simulation set to experiments performed at different CHOL concentrations. We also evaluated whether the simulation models reproduce the dependence of lateral diffusion coefficients on cholesterol measured using pulsed field gradient NMR \cite{filippov2003effect,filippov2003influence}. This comparison is enabled by the recent theoretical work that allows for the elimination of finite-size effects in MD simulations and thus a quantitative comparison with experiment \cite{vogele2016divergent,vogele2018hydrodynamics}. With the structural and dynamic comparisons established, we estimated the quality of each force field at different CHOL concentrations, identified the best models, and analyzed their molecular organization as we expect them to best reflect reality.

@@

\subsection{X-ray Scattering Experiments}
\todo{Scattering methods incomplete}

Such scattering experiments are challenging to interpret for multi-component lipid bilayers \cite{pan12,Heftberger15,Marquardt15}, yet we still extract 
thus area per molecule values, one of the main quantity used to
compare simulations to experiments, have not been
available for lipid cholesterol mixtures.

SAXS data on POPC multilamellar vesicles (MLVs) at various CHOL concentrations has been measured. Data have been obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons, $T=27^\circ$C. Data were analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}. Data from MLVs are a convolute of structure factor (the crystalline lattice) and form factor. By fitting the scattered intensity data we obtain both contributions. Here we posted only form factors (ASCII format). For information on the quality of the fit we also give plots of the fitted intensity data. The electron density profile has been modelled in terms of the SDP model \cite{heberle12,Kucerka08a,kucerka12}, that is volume distribution functions are modelled by individual Gaussians or error functions. CHOL is challenging to describe within the SDP model, and instead of density profiles, the form factors are often compared between simulation and experiment \cite{Kucerka08b}, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}. Here, cholesterol is accounted for by two Gaussians, as proposed by Jianjun Pan (private communication), but is to the best of our knowledge not published.
%(see also PhD Thesis by Peter Heftberger). 
Additional figures show the volume distribution functions and the resulting electron density profiles.

@@

\subsection{Molecular Dynamics Simulations}

We performed MD simulations of a pure POPC membrane as well as five POPC/cholesterol mixtures with the cholesterol content ranging from 11 to 47~mol-\%. In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes. The small (64 POPC) and medium (256 POPC) systems were only used in the diffusion analysis, whereas all other analyses were performed on the large membranes (1024 POPC). The number of POPC molecules was kept constant across the different cholesterol concentrations. All membranes were solvated by 50 waters per lipid (POPC or cholesterol). The simulations were performed using four commonly used force fields, namely CHARMM36 \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig15b} models. All simulations were 1~\textmu{}s long, totaling 72~\textmu{}s, and performed using GROMACS version 2020 or 2021 \cite{pall2020heterogeneous}. The used simulation parameters are provided in Table~\ref{SItab:params}. The simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog).

\begin{table*}[]
\begin{center}
    \caption{\label{tab:simulations}%
    Details of the simulation systems provided for (small/medium/large) system sizes). The box dimensions in the membrane plane ($x/y$) and normal to the membrane ($z$) are provided for the Slipids simulations, and the values vary slightly between the force fields.
    }
    \begin{tabular}{c|ccccc}
    \toprule
    Cholesterol & POPC & Cholesterol & Water & $x/y$ (nm) & $z$ (nm) \\
    \midrule
    0~mol-\%    & 64/256/1024 & 0/0/0       &   3200/12800/51200 & 4.4/9.0/18.1 & 8.9/8.6/8.5    \\
    11~mol-\%   & 64/256/1024 & 8/32/128    &   3600/14400/57600 & 4.5/9.4/18.1 & 9.4/9.1/9.3    \\
    20~mol-\%   & 64/256/1024 & 16/64/256   &   4000/16000/64000 & 4.5/9.2/18.3 & 10.2/9.9/10.0  \\
    29~mol-\%   & 64/256/1024 & 26/104/416  &   4500/18000/72000 & 4.6/9.2/18.5 & 10.8/10.8/10.7 \\
    38~mol-\%   & 64/256/1024 & 40/160/640  &   5200/20800/83200 & 4.8/9.5/19.2 & 11.3/11.4/11.2 \\
    47~mol-\%   & 64/256/1024 & 56/224/896  &   6000/24000/96000 & 5.0/10.1/20.0 & 11.8/11.6/11.7 \\
    \bottomrule
    \end{tabular}
\end{center}
\end{table*}

\subsection{Simulation Analyses}

\paragraph{Electron Density Profiles and Bilayer Dimensions:} The electron density profiles were calculated using \texttt{gmx density} after assigning all atom types with valence electron counts. The thicknesses were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data. Area per phospholipid was obtained by dividing the area of the bilayer by the number of phospholipids in one leaflet.

\paragraph{Lateral diffusion coefficients:} The lateral diffusion coefficients $D_\mathrm{PBC}$ from simulations performed using periodic boundary conditions (PBC) were extracted from mean squared displacement (MSD) data calculated for lipid centers of mass after eliminating the drift of their host leaflet. The MSD data were fit with a straight line in the lag time ($\Delta$) interval between 10 and 100~ns as
%
\begin{align}
	\mathrm{MSD}=4D_\mathrm{PBC}\Delta.
\end{align}

The diffusion coefficients extracted from the three simulation box sizes were fitted with
%
\begin{align}
	D_\mathrm{PBC}\approx D_\infty+\frac{k_\mathrm{B}T}{4\pi\mu_\mathrm{m}h}\frac{\ln\left[L/\left(L_\mathrm{SD}+1.565H\right)\right]-1.713}{1+H/L_\mathrm{SD}},
	\label{eq:sdlatpbc}
\end{align}
%
where $D_\infty$ is the lateral diffusion coefficient in an infinite system, $h$ is the hydrodynamic thickness of the membrane, $k_\mathrm{B}T$ is the Boltzmann constant, $T$ is the temperature, $mu_\mathrm{m}$ and $mu_\mathrm{f}$ are shear viscosities of the membrane and the fluid (water), respectively. $H$ is half the thickness of the water layer and $L_\mathrm{SD}=\frac{h\mu_\mathrm{m}}{2\mu_\mathrm{f}}$ the Saffmann--Delbr\"{u}ck length \cite{vogele2018hydrodynamics}. The inter-leaflet friction coefficient does not appear in Eq.~\eqref{eq:sdlatpbc} as we expect it to be infinite, which was found to be a valid assumption for lipid bilayers \cite{vogele2018hydrodynamics}. The water viscosity value of 0.3228~mPa$\cdot$s was interpolated to 298~K from the values for CHARMM TIP3P in Ref.~\citenum{ong2019temperature} and used for all simulations (CHARMM TIP3P and normal TIP3P differ by $\sim$2--3\%).

\paragraph{More and Less Ordered Populations:} The angles ($\theta$) between the C--H vectors of the last six carbons of the palmitate chain were extracted and the data across all CHOL concentrations was accumulated. These data were fit with two Gaussians centered at 0,
%
\begin{align}
    P(\theta)=a_\mathrm{less}\exp\left(-\frac{\theta}{2\sigma_\mathrm{less}}\right)+a_\mathrm{more}\exp\left(-\frac{\theta}{2\sigma_\mathrm{more}}\right),
    \label{eq:twogauss}
\end{align}
%
where the two standard deviations ($\sigma$) correspond to the more and less ordered components. After these $\sigma$ values were established for a given force field, the data at all CHOL concentrations was independently fitted by Eq.~\eqref{eq:twogauss} with fixed $\sigma$ values. These fits provided the pre-factors, whose ratio $a_\mathrm{more}/(a_\mathrm{more}+a_\mathrm{less})$ corresponds to the fraction of the more ordered component in the system.

This two-Gaussian approach assumes a ``two-population'' scheme, where the more and less ordered populations are present (at non-zero CHOL concentration), and that the addition of CHOL only affects their ratio, but not the nature. In an alternative picture of a ``collective transition'', there would be a single Gaussian, whose standard deviation $\sigma$ would vary systematically as a function of CHOL concentration.

\paragraph{Quantitative Quality Measure for Lipid Bilayer Structure:} The deviations (in \%) from experimental values across CHOL concentrations were used to quantify the quality of the lipid force fields. The density profiles (as a function of distance from membrane core along the membrane normal), oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains), and diffusion coefficients were interpolated to create continuous 2D matrix for density profiles and order parameters, and a continuous 1D vector for the diffusion coefficients. For the 2D matrices, the absolute values of the differences of the matrices from simulations and experiments were calculated and averaged over one dimension which was either the distance from membrane core or the carbon atom number, respectively. This resulted in a 1D deviation vector as a function of CHOL concentration. For diffusion coefficients, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration. All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \%.

\todo{Results: volmaps of chains around cholesterol?}\\
\todo{Results: SI: RDFs of POPC/CHOL around CHOL}\\
\todo{Results: Add missing 30\% data to the scattering plots}\\
\todo{FF minima analysis}

\subsection{Cholesterol Condensing Effect is Manifested Differently in Different Force Fields}

Lipid bilayer dimensions can be experimentally accessed by measuring X-ray scattering form factor, which is related to the electron density along membrane normal \textit{via} a Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16,??}. The form factor can be translated to an electron density profile, and area per lipid and bilayer thickness values can be extracted using the scattering density profile (SDP) model or its combination with MD simulations  \cite{pan12,Heftberger15,Marquardt15,??,doktorova2020molecular}. Sophisticated approaches for single component lipid bilayers are available but the data for multi-component systems are more difficult to interpret, and the used approaches involve severe approximations \cite{pan12,Heftberger15,Marquardt15,??}.

Thus, we compare here the total electron density profiles which do not require the de-convolution of the electron density profiles into densities of the different molecular species. For reference, all form factor profiles from experiment and simulation are shown in Fig.~\ref{SIfig:scattering}. As the absolute and relative lobe amplitudes depend on simulation box size (Fig.~\ref{SIfig:scatteringvssize}), the minima in the form factor profiles provide a robust comparison between simulation and experiment. Thus, we have also traced the minima in these profiles, and they are plotted in Fig.~\ref{SIfig:ffminima}. As simulations and experiments were performed at somewhat different CHOL concentrations, we generated two-dimensional maps for the density profiles using interpolation (see Methods). These maps are shown in Fig.~\ref{fig:densmaps} for both simulations (top row) and experiment (bottom row, left column). The original electron density profiles used in the generation of these maps are shown in Fig.~\ref{SIfig:densprofs}.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\linewidth]{../FIGS/densitymaps.pdf}
  \caption{\label{fig:densmaps}%
  \textbf{Electron density profiles, thickness and area per lipid as a function of CHOL concentration.}
  %
  Top: Electron density maps for the simulations using four different force fields.
  %
  Bottom left: Electron density map for the experiment. The scaling is uniform between all simulations and experiment. The original electron density profiles are shown in Fig.~\ref{SIfig:densprofs}. The scattering experiments are interpolated to 298~K from Refs.~\citenum{pabst2000structural} (58.8~\AA{}$^2$), \citenum{kuvcerka2006structure} (67.5~\AA{}$^2$), and \citenum{Kucerka11} (63.5~\AA{}$^2$). The NMR data is measured in Ref.~\citenum{Leftin14} using either $^{13}$C (58.9~\AA{}$^2$) or $^2$H (58.3~\AA{}$^2$) NMR. The effect of system size on the density profiles in simulations is demonstrated in Fig.~\ref{SIfig:densprofssize} in the SI.
  %
  Bottom center: Bilayer thickness. Thickness is defined as twice the distance from the peak in electron density to the membrane core. Experimental data are extracted in a similar manner from electron density profiles obtained with X-ray scattering.
  %
  Bottom right: area per lipid measured by dividing the total membrane area by the number of phospholipids. 
  %
  The size-dependency of area per lipid is shown in Fig.~\ref{SIfig:aplvssize}.
  }
\end{figure}

Overall, all electron density profiles share the same features across all CHOL concentrations; a high-density band corresponding to the tightly-packed interfacial region rich in electron-rich phoshorus, a low-density region at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase. However, a more detailed look at the profiles in Fig.~\ref{fig:densmaps} reveals differences between the force fields. CHOL renders the membrane thicker, resulting in the deviation of the high-density bands with increasing amount of CHOL. In CHARMM36, Slipids, and MacRog, there is clear saturation of this effect at around 30~mol-\% CHOL. In Lipid17, however, the membrane gets thicker until the maximum studied CHOL concentration of 47~mol-\%. In experiment, the effect of CHOL is generally weaker, and some saturation is evident at high CHOL concentrations. These trends are visible in the bottom middle panel in Fig.~\ref{fig:densmaps}, which shows the effect of CHOL on membrane thickness. Fig.~\ref{fig:densmaps} also demonstrates that CHARMM36 has the sharpest low- and high-density bands indicating, and the same is true for MacRog at higher cholesterol concentrations. This uniform membrane structures could indicate small undulations, whereas the less sharp Lipid17 and especially Slipids profiles could indicate a more flexible membrane.

Comparison against experiments (bottom left panel in Fig.~\ref{fig:densmaps}) indicates that the membrane structure is more uniform in the experiment, as the bands are sharper. This is also evident in the original electron density profiles in Fig.~\ref{SIfig:densprofs}. System size plays a role here, as demonstrated by the density maps calculated for the CHARMM36 simulations in three sizes, shown in Fig.~\ref{SIfig:densprofssize}. The larger the system, the larger the bilayer undulations are, which can smear the features of the density profiles. Still, even in the smallest system, the band intensities are not as localized as in the experiment.

Comparison of the thickness values (bottom middle panel in Fig.~\ref{fig:densmaps}) demonstrates that the Lipid17 bilayers are too thin at low CHOL concentrations, whereas other models agree well with experiment in this regime. Still, all simulation models overshoot the experiment at high CHOL concentrations. Experiment and all simulation models except Lipid17 demonstrate saturation after 30~mol-\%.

The membrane-condensing effect of CHOL is well documented, yet there is some ambiguity surrounding the term \cite{edholm2005areas}. The cross-sectional area of a CHOL molecule is smaller than that of a phospholipid, so an increase in CHOL concentration is expected to lower the average \emph{area per lipid} (including CHOL). However, condensation could also refer to a decrease in the \emph{area per phospholipid} upon the addition of cholesterol \cite{edholm2005areas}, or to CHOL having a diminishing partial area, \textit{i.e.} a certain amount of CHOL could be added to a phospholipid bilayer without increasing or decreasing its total area \cite{javanainen2017two}. We thus compared the behavior of the area per (POPC) lipid (APL) as a function of CHOL concentration in the bottom right panel of Fig.~\ref{fig:densmaps}. The dependence of APL on CHOL concentration follows the trends in thickness inversely in general, yet provide curious differences between the force fields. Lipid17 has the largest APL across the entire CHOL contration range, and the addition of CHOL always leads to an increase in the membrane area. MacRog also has a large APL for pure POPC, but the partial area of CHOL is negative until 30~mol-\% concentration indicating a particularly strong condensation effect. The profiles for Slipids and CHARMM36 are very similar, and the partial area of CHOL is small or even zero until a concentration of 20~mol-\%. Importantly, the observed differences in CHOL condensing effect take place at the physiologically relevant CHOL concentration range \cite{van2008membrane}. Experimental data are derived from models, so they can vary greatly, as evidenced by the very different values obtained by \citeauthor{pabst2000structural}\cite{pabst2000structural} and \citeauthor{kuvcerka2006structure}\cite{kuvcerka2006structure} for POPC at 303~K.

\subsection{Acyl Chain Ordering Varies Greatly Between the Force Fields}

Order parameters for C--H bond vectors measured with with $^{13}$C or $^{2}$H NMR techniques, provide indirect information about the structural sampling of indidual molecules \cite{ollila16}. Moreover, these parameters are readily extracted from the simulation. An increase in acyl chain order parameters with added cholesterol in simulations and experiments can be explained by increased anti conformations in acyl chains \cite{ferreira13,??}, which is suggested to play critical role in the phase behaviour of PC--cholesterol mixtures \cite{ipsen87}. Consequently, the correct cholesterol ordering effect is expected to be a necessary condition for a model used to understand lipid--cholesterol phase behaviour.
